Clustering galaxies at 0.5 < $z$ < 1: an unsupervised approach

### 2017-06-17 main_all_together.tex
Moving around text, working on ref letter

@@

    For each generated network, we compared the results with K96 categorization.
    We also used these networks to classify the T12 galaxies samples, and compared them with classifications from supervised networks in T12.
    \subsection{1D SOMs}
    \label{sec: 1D}
        \subsubsection{TRAINING THE NETWORKS}
        \label{sec: 1Dt}
            To start our clustering, we assumed that galaxies can be divided only in two types; quiescent and starburst.
            In this case, we generated a network with only two neurons.
            We increased the size of the map gradually till we separated the input 12 sample galaxies into the 12 neurons.

            Figs.~\ref{fig: 1by2T} -~\ref{fig: 1by22T} show results of training networks.
            Similar to Fig.~\ref{fig: sample}, the upper part of the figures shows neurons and their relative distances, $D_j$, from each other.
            As it was mentioned in the Sec.~\ref{sec: method}, the colours between neurons represents the relative distance between the neurons and the darker is the colour the more relative distance is between them.
            The lower part shows how many galaxies from K96 template places in each neuron. 
            \begin{figure}
                \begin{subfigure}[b]{0.5\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{../images0.01/1d/dist_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.5\textwidth}
                     \includegraphics[width=\textwidth]{../images0.01/1d/hit_t_1_by_2.png}
                \end{subfigure}
                \caption{Results of training network in $1\times2$~grid. Similar to Fig.~\ref{fig: sample} the upper plot is a distance map and the lower plot is a hit map. In this network, 5 of galaxies from K96 model, categorized as quiescent galaxies and the rest are starbursts. Because of their high emission lines, Sc galaxies moved towards the starburst ones.}
                 \label{fig: 1by2T}
            \end{figure}

@@

            The fact that the galaxies had to have 22 different neurons till be divided into 12 groups, shows that the difference between SB1 and SB2, and B and E galaxies are very small.
            These small differences might be very small.
            Therefore, they intend to stay in the same group till the network become big enough to make attention to the smallest particularity.

>>>>

\documentclass[useAMS,usenatbib]{mn2e}
\bibliographystyle{mn2e}

\usepackage{amsmath}
\usepackage{commath}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{times}
\usepackage{float}
%\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{color,soul}
\usepackage{import}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{amssymb}	% Extra maths symbols
\usepackage{multicol}        % Multi-column entries in tables
\usepackage{bm}		% Bold maths symbols, including upright Greek
\usepackage{pdflscape}	% Landscape pages
%\usepackage{booktabs,fixltx2e}
\usepackage[flushleft]{threeparttable}

\newcommand \Spitzer {{\it Spitzer }}
\newcommand{\boldit}[1]{\textbf{\mathversion{bold}#1}}

\newcommand \aaj {A\&A}
\newcommand \aarv {A\&ARv}%: Astronomy and Astrophysics Review (the)
\newcommand \aas{A\&AS}%: Astronomy and Astrophysics Supplement Series
\newcommand \afz {Afz}%: Astrofizika
\newcommand \aj {AJ}%: Astronomical Journal (the)
\newcommand \apss {Ap\&SS}%: Astrophysics and Space Science
\newcommand \apj {ApJ}
\newcommand \apjs {ApJS}%: Astrophysical Journal Supplement Series (the)
\newcommand \araa {ARA\&A} %: Annual Review of Astronomy and Astrophysics
\newcommand \asp {ASP Conf. Ser.}%: Astronomy Society of the Pacific Conference Series
\newcommand \azh {Azh}%: Astronomicheskij Zhurnal
\newcommand \baas {BAAS}%: Bulletin of the American Astronomical Society
\newcommand \mem {Mem. RAS}%: Memoirs of the Royal Astronomical Society
\newcommand \mnassa {MNASSA}%: Monthly Notes of the Astronomical Society of Southern Africa
\newcommand \mnras {MNRAS} %: Monthly Notices of the Royal Astronomical Society
%\newcommand {Nature}%(do not abbreviate)
\newcommand \pasj {PASJ}%: Publications of the Astronomical Society of Japan
\newcommand \pasp {PASP}%: Publications of the Astronomical Society of the Pacific
\newcommand \qjras {QJRAS}%: Quarterly Journal of the Royal Astronomical Society
\newcommand \mex {Rev. Mex. Astron. Astrofis.}%: Revista Mexicana de Astronomia y Astrofisica
%\newcommand {Science }%}%(do not abbreviate)
\newcommand \sva {SvA}%: Soviet Astronomy
\newcommand \aap {APP} %:American Academy of Pediatrics
\newcommand \apjl {ApJL} %:The Astrophysical Journal Letters

%\defcitealias{Hossein12}{T12}

\begin{document}
% TITLE
\defcitealias{Hossein12}{T12}

\title[SOM: classifying high $Z$ galaxies]{Clustering galaxies at 0.5 < $z$ < 1: an unsupervised approach} %PB160426: suggest "Clustering galaxy spectra at $0.5<z<1$: an unsupervised approach" -- I think if you say "clustering galaxies" people will think of spatial clustering
\author{rahmani.sahar }
\date{\today}
\author[S.~Rahmani, H.~Teimoorinia and P.~Barmby]{S.~Rahmani$^{1}$\thanks{E-mail:
srahma49@uwo.ca}, H.~Teimoorinia$^{2}$, P.~Barmby$^{1}$\\
$^{1}$Department of Physics $\&$ Astronomy, Western University, London, ON N6A 3K7, Canada\\
$^{2}$Department of Physics $\&$ Astronomy, University of Victoria, Finnerty Road, Victoria, British Columbia, V8P 1A1, Canada}
\maketitle

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%abstract
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\begin{abstract}%% (Must be 250 Words)
% In this paper, we introduce a new method of clustering of high red-shift galaxies, based on their spectral energy distributions (SED)s. We use unsupervised self organizing map(SOM) to train an artificial neural network to categorize SEDs of galaxies. We utilize as spectral template as a training set, and trained networks to be able to clustered galaxies based on the template. We cluster a sample of 142 galaxies with 0.5< z <1 using the trained network. We train networks with various sizes and properties to investigate the nature of the 142 sample galaxies. For each network, we cluster the sample of 142 galaxies and visualise their properties. In order to test new classifications,we compare physical properties of the galaxies in each cluster. We compare our results with  clustering  results  from  other  artificial  neural  networks,  and  with  the  original template classification. We conclude that SOM is a very useful method to be used in classifying galaxies based on their SEDs. It can be used to find morphology of the high red-shift galaxies, and find a new class of SEDs.
In this paper, we introduce a new method of clustering high red-shift galaxies, based on their spectral energy distributions (SED)s.
%Artificial neural networks provide us a reliable method to cluster SEDs of galaxies.
We use an unsupervised self organizing map(SOM) as a training method to train networks with a sample of galaxies with SEDs between FUV to NIR wavelengths, which their morphological types are known.
Since information that can be extracted from each network depends on the size and dimension of the trained networks, in this project we train networks with different grids in 1D or 2D.%%???
We visualise each trained network using a new version of SOM plots.
To test trained networks, we cluster a sample of 142 galaxies with 0.5 < $z$ < 1 using each of the trained networks.
We plot average SED of SEDs of galaxies in each neuron in 1D networks, to investigate the nature of galaxies in new clusters.
For each network, we visualise the properties of galaxies in each neurons. %%%% combine 94 and 93 
We compare mean physical properties of the galaxies in each neuron such as age, specific star formation rate, stellar mass, and FUV extinction, with each other and found a tight correlations between them. 
Comparing our results with clustering results from supervised artificial neural networks, show us using SOM we can classify galaxies in between types of original spectral types. %%rephrase
As a result we have a power tool to classify spectra of galaxies based on their SEDs which also can  identify galaxies with a new shape of SEDs.

%%%% SR20160512: I do not like it! But I do not know how to fix it. I already wrote at least 4 version of it. I need help with this one:((

%%% what is the problem we are trying to solve in general! why do we want o classify thing in general, and what is wrong in exciting classification scheme

\end{abstract}
\begin{keywords} 
 galaxies: high red shifts, galaxies: spectral energy distribution, methods: observational, methods: statistical, data mining, methods:data analysis
\end{keywords}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Intro
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Introduction}
\label{sec: intro}
%General information about SEDs
All information we can obtain from a galaxy is encapsulated in the light it emits; every observable phenomenon in a galaxy leaves a footprint on the spectral energy distribution (SED) of that galaxy.
We can determine some physical properties of galaxies by properly modeling various features observed in their SEDs.
The general shape of the SEDs can be used as an identifier of the morphological type of the galaxies.
Considering the main features of the SEDs galaxies can be categorized in two main groups: elliptical, or spiral.
Each group has its own characteristic features and can be divided into many sub-branches.

%%Creating templates and NIR_UV templates
Several attempts have been made to composite a thorough template for categorizing the spectral type of galaxies using data from nearby galaxies~\citep[e.g.][]{Kinney93}~\citep[][hereafter K96]{Kinney96}~\citep[][]{Bershady00,Mannucci01}. 
Based on their usage, these templates are restricted to certain wavelengths.
In near-infrared (NIR) to ultraviolet (UV) domain of wavelengths (the region where output of stars peaks), SED contains information about the main physical properties of galaxies, e.g., age, star formation rate (SFR), stellar mass, wide range of the stellar population, some information on interstellar medium (ISM)'s absorption and emission lines, and extinction from ISM of the galaxies.

% categorizing galaxies based on their SED

Attaining high-resolution data and detailed SEDs has been made possible due to advances in imaging techniques, and photometry and spectroscopy devices.
This has helped to see the more detailed SEDs, which made classification of galaxies more complex.
Since no two galaxies, even with the same morphology, have exactly the same properties, classifying the SED of the galaxies using the templates are a challenge.
Many fitting methods are developed and used to find the best matches of template for each SED.
$\chi^2$ minimizing method is the most commonly used method to do so. 
Artificial neural networks (ANNs), K-mean clustering, and principal component analysis are the other methods to cluster and classifying the morphological type of the galaxies based on their SED \citep[e.g.][]{Allen13,Ordov14,Shi15}.

%ANNs
ANNs are very powerful tools to use in data processing and pattern recognition problems.
Their information processing method was inspired by the way neurons in human brains work.
An ANN contains many interconnected units (nodes or neurons) which process data and work together to solve problems.
It uses a set of training methods to learn about a non-linear and complex relations between input and output data and how to apply it to new sets of data.
Studies show that ANNs provide better fitting than minimizing chi-squared and would be an alternative choice for fitting data~\citep[e.g.][]{Marquez91,Moayed09}.
In cases that both methods show the same correlation, ANNs still are a better method of fitting due to their faster results on large databases \citep[][]{Gulati97}.

%Training methods for networks
Neural networks can be trained using two methods; supervised and unsupervised.
In supervised method, a neural network would be trained using input data based on desired outcome.
This method is very useful for classification of data with specific target values (e.g. any pattern recognition with known templates).
On the other hand, in unsupervised method there is no prediction of output data.
This method classifies data based on their underlying structures and hidden variables.
The unsupervised method is a very helpful method to knowledge discovery of your data, or when the underlying structure of data is not well established (e.g. producing a template of SED of galaxies).

%SOM
Kohonen Self organizing map (or self organizing map, SOM) is an unsupervised neural network for mapping and visualizing a complex and non-linear high dimension data introduced by~\citep{Kohonen82}.
The SOM is a technique that shows a simple geometry relationship of a non-linear high dimension data on a map \citep{Kohonen98}.
%SOM in Astronomy
The utilization of the SOM in astronomy dates back to 1990s. 
\citet[][]{Odewahn92}, \citet[][]{Hernandez94}, and \citet[][]{Murtagh95} were among the first studies in astronomy that used SOM for their studies.
From classifying quasars' spectra to star/galaxy classifications, from gamma-ray bursts clustering to classification of light curves, this method have been used in various aspects in astronomy \citep[e.g.][]{Maehoenen95, Miller96,Andreon00,Balastegui01,Rajaniemi02,Brett04,Scaringi09}.
%20160509PB: suggest starting a new paragraph here, with an introductory sentence about galaxy spectra.
%SR20160512: I start new paragraph after the next sentence not here.
\citet{Geach12} used COSMOS data to demonstrate two of main applications of SOMs; object classification and clustering, and photometric redshift estimation. 
The later one was subject of many studies \citep[e.g.][]{Kind14a}.

%SOM in Astronomy continued.
Spectrum of a galaxy is a combination of spectra of million stars and clouds inside the galaxy.
Therefore, large spectra data sets contain information of billions of stars and other objects inside galaxies, and needs to be studied carefully. %%%EEEHHHH
\citet{In12} introduced a new clustering tool, which is based on the SOM method for analysing large data of spectra.
They used $\sim 60000$ spectra from the Sloan Digital Sky Survey \citep[SDSS;][]{Abazajian09}
to test their tool, and created very large SOMs to analyse the type of spectra/objects.
They also generated SOMs from quasars' spectra in order to find unusual types of quasars' spectra. Later \citet{Meusinger16} used these SOMs and updated data from SDSS and other surveys and found a new class of quasars.
The other application of the SOM is to find outliers or errors in the data.
\citet{Fustes13} produced a package based on SOM to classify spectra from GAIA survey that were previously classified as "unknown" by SDSS pipeline. The package recognizes an astronomical object from artificial errors, and then classifies the object based on its spectra.

%What T12 did
~\citet[][hereafter T12]{Hossein12} classified SEDs of this sample of galaxies using supervised neural network method, based on the spectral template presented by K96.
With supervised method, they could only classify SED of the 105 out of 142 galaxies.
The SED of the 37 galaxies from their samples could not be matched with any of those in the K96 templates. 
In order to classify the remaining 37 galaxies, they combined spectrum of the K96 templates.
This procedure showed that, not a single type of galaxies in K96 template could describe the SED of any of those 37 galaxies.
Since the SOM have the freedom of classifying objects in between the known classes, it be applied on this data to find the best SED class for the 37 galaxies.
T12 also showed that there are tight correlations between physical properties of galaxies and these correlations might be different for each type of the galaxies.

%What we did
In order to compare supervised and unsupervised methods directly, we used exactly the same data as T12 used.  
First, we trained SOMs using data from K96 and compared our classification with K96 galaxies classification.
Then, we used the trained network to classify the SEDs of 142 galaxies with 0.5 < $z$ < 1 from T12 paper.
Same as T12, We also plotted the properties of the T12 galaxies based on the new clusters and compared them with previuse works.
In Section $\S$~\ref{sec: data}, we present the data that we used to train and test our networks. We describe the SOM methos in Section $\S$~\ref{sec: method}. The results of the SED classifications and comparing them with previous studies are shown in Section $\S$~\ref{sec: result}. In Section $\S$~\ref{sec: summary}, we present the summary of our results and the future works in this subject.

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%DATA
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\section{DATA}
\label{sec: data}
We used SED templates from K96, to train neural networks.
To test the trained networks, we used SED and physical properties of 142 galaxies at 0.5 < $z$ < 1 from T12.
Following of T12 work, we chose these two sets of data to not only show the application of SOMs in SED clustering, but also to easily compare both supervised and unsupervised methods.

 \subsection{Kinney spectral model}
     \begin{figure}
        \centering
        \includegraphics[width=0.5\textwidth]{images/k96.jpg}
        \caption{K96 spectra template for 12 types of galaxies from T12 paper (Fig. 1 in T12).Type of each template is shown in each frame. Plots B, E, S0, Sa, Sb and SC show spectra that belongs to early type galaxies. Starburst galaxies spectra are indicated with SB 1 to 6. Higher numbers represent more intrinsic colour extinction.}
        \label{fig: k96}
    \end{figure}

    K96 used UV-optical spectra of 70 star forming and quiescent nearby galaxies to produce a template that contains 12 types of SEDs.
    These templates are widely used in many studies to determine morphological type of galaxies, or find properties of specific types of galaxies in many studies \citep[e.g.][]{Shakouri16,Paiano16,Laporte16,Holden16}.
    K96 mentioned that one of the usage of these templates can be classifying the SED of the high redshift galaxies. 
    The 12 types of spectra are divided based on their morphological types for early type galaxies or their extinction for starburst galaxies (Fig.~\ref{fig: k96}).

    The quiescent group of galaxies includes Bulge (B), Elliptical (E), S0, Sa, Sb, and Sc galaxies.
    Bulge group represents galaxies similar to M31 and M81, which their UV and optical spectroscopy dominated by the stellar population in their bulges.

    The starburst galaxies are divided into six groups (SB1 - SB6) based on their intrinsic extinctions ($E(B-V)$). 
    As it is clear in the Fig.~\ref{fig: k96}, SB1 galaxies have lower internal extinctions ($E(B-V) \simeq 0.05$), while SB6 galaxies have the highest amount of extinction ($E(B-V) \simeq 0.65$) among starburst galaxies.

    K96 spectra spans from $\sim1200-10000$\AA with resolution of $\sim 10$\AA~.
    However, we only used data between $\sim1200-8000$\AA~to train our networks. 
    This wavelength range was chosen due to availability of flux information in those wavelengths for all 12 types of galaxies.
    In early type galaxies' spectra (B to Sb), the spectrum is redder and strong absorption lines, and 4000\AA~break is distinguishable. 
    While SEDs of starburst galaxies are more flat on the optical and NIR part than early type one and show strong emission lines.
    For more details on each spectral type, we encourage readers to see K96 and references therein.

 \subsection{SED and Properties of the sample galaxies} 
    T12 selected 142 galaxies from the spectroscopic campaign of the ESO GOODS-South field.
    For each galaxy, robust spectroscopic redshift and photometry data from the Good-MUSIC catalogue had been known.
    The 142 galaxies had been selected in T12 due to availability of data from HST/ACS; VLA/ISAAC; and \Spitzer/MIPS and IRAC for whole galaxies. 
    Data from these instruments were necessary in order to have a complete picture of stellar population and SFR.
    The photometry data contain information of the galaxies in 10 - 13 filters with the wavelength range of $\sim 0.4-24 \mu$m in the observed frame.
    T12 matched the resolution of the photometry data with K96 data and used these photometry data as inputs for Code Investigating GALaxy Enmission ({\em CIGALE} code;~\citep[][hereafter N09]{Noll09}) to generate the best fitted SED for each galaxy.% as well as some of the physical properties of the galaxies.

    {\em CIGALE} is a valuable tool to investigate the properties of the galaxy using UV to IR wavebands.
    It uses stellar populations, synthetic attenuation and dust emission models, spectral line templates, and active galactic nuclei's optical spectral templates to model SED of the galaxies.
    This code was successfully tested on data from 39 galaxies selected from the Spitzer Infrared Nearby Galaxy Survey (SINGS;~\citep{Kennicutt03}) by N09.
    T12 produced the best SED match, with wavelength interval of 910 to $\sim 8E9$\AA~, for each galaxy.
    Assuming decreasing SFR and visual attenuation ($\tau$) model, Salpeter initial mass function~\citep{Salpeter55} and old stellar population with age of $\sim 10$~Gyr, they derived physical properties of galaxies such as age, and stellar mass.
    Some of these properties are shown in Tab~\ref{tab: props}.
    In the Sec.~\ref{sec: 1D}, we studied these properties for each categorization.
    More details on creating SEDs and extracting information about galaxies properties using {\em CIGALE} can be found in N09 and T12.

    \begin{table}
\caption[]{Description of the properties of T12 galaxies; the output result of {\em CIGALE}}     
\label{tab: props}
\centering
\begin{tabular}{l l l}
\hline\hline
\noalign{\smallskip}
Par. & Unit & Description\\
\noalign{\smallskip}
\hline
\noalign{\smallskip}
$t_{\,\mathrm{oSP}}$ & Gyr & age of old SP model \\
$t_{\,\mathrm{ySP}}$ & Gyr & age of young SP model \\
$f_\mathrm{burst}$ & --- & mass fraction of \\
& & young single population (SP) model \\
\noalign{\smallskip}
$t_{\,\mathrm{D4000}}$ & Gyr & D4000-related age \\
\noalign{\smallskip}
$M_\mathrm{star}$ & M$_\odot$ & total stellar mass  \\
SFR & M$_\odot$/yr & instantaneous SFR  \\
$A_\mathrm{FUV}$ & mag & attenuation at 1500\,\AA{} \\
\noalign{\smallskip}
\hline
\end{tabular}
\end{table}

    For testing the created networks, we used SEDs that were produced by T12. 
    These SEDs are publicly available in the form of flux per rest frame wavelength in a wide range of wavelengths.
    Since we have used the T12 SEDs to test trained network, we only used the part of the SEDs that have a same wavelength range as K96 templates.

 %%%%SR160611: The following is the Hossein's answers to the questions that you asked in the data section; Since, we did not used the spectra data here (we only used generated SEDs and information that was derived from SEDs) ,I assumed it is unnecessary to talk a lot about spectroscopy data. But, if you think it would help to understand the subject better, I will add those information to data section.

%  Is the data vector for each object a list of fluxes (flux densities) as a function of (rest) wavelength? I think we discus this question so I skip it.
%  Same wavelengths for every object? Yes
%  What is the spectral resolution? It depends on the type of spectra we use in T12 see
% below for more information.
%  Do the spectra include the whole galaxy or just the central part? For observed 142
% galaxies: whole spectra because they are in high redshift. For K96 they are co-added
% spectra from different parts.
%  Does the resolution match with K96? We match them to K96, see below.
%  What is the coverage of spectra? See below.
%  Do these galaxies have spectroscopic redshifts, or were redshifts derived from
%   photometry?" All 142 galaxies have robust spectroscopic redshift.
% For more clarification about spectra used in T12: We have three sets of spectra.
% 1 – observed spectroscopic spectra. They are used just for extra independent validation; see figure 17 from T12. This validation is nothing to do with network validations. The useful wavelength range (of
% FORS2 spectrometer) is 4800−10 000 Å, the nominal resolution is R = λ/Δλ = 580, which corresponds to
% a spectral resolution of ∼13Å. We use only range [3500-5100] because all the 142 observed spectra have information in this interval (some of them have no spectral information in 8000A, for example). In T12 you can also find this that can help: “The rest-frame spectra of the sample all include spectral information
% ∼3500–5100 Å. This interval brackets the important features such as the 4000Å break and also the strong emission
% lines such as [O ii]λ3727, λHβ 4861, and [O iii] λλ4959, 5007. The spectroscopic data for the sample galaxies contain different spectral types from starbursts (with a flatter continuum and stronger emission lines) to early-type galaxies (with a redder continuum, strong absorption lines, and large 4000Å break.”
% 2- k96 Model spectra which are for training. The spectra cover the interval ∼1200–10000Å and have a resolution of∼10 Å. We use only interval 1200-7500A. because all the 12 K96 spectra have information in
% this interval (some of the spectra have no information in 9000A, for example).
% 3– The Maraston’s models (SEDs): For each galaxy (from 142) we obtain a best Maraston spectrum (SED) by fitting photometric data to the models. Mraston’s models range very extended wavelength interval (from UV to far IR) with different resolution in different part of a given spectrum. For example resolution in IR is 100A and in optical ~3-5A. The best resolution is related to the optical part (3-5A and we convert all to 10A i.e., to resolution of K69). We cut the best 142 spectra in interval [1200-7500] like K96 for ANN validation or test. (If we had spectral information for all K96 spectra in the interval 1200-10000A, for example, then we would cut all 142 SEDs in 1200- 10000A). So if you find a new training set (other than K96 which we talked about that in our skype chat) and if they had information in 1000-15000A, for example, then we would cut the SEDs in this new interval for validation.
% As a summary: as you know (better than me!) we use K96 as training set and 142 SEDs as validation (or test) set. When we obtain the class of each 142 galaxies (in fact 105 galaxies! You are improving this number to 142 in your paper) then we use 105 observed spectra to make figure 17 from T12. The co-added observed spectra in a certain class match with the associated K96 spectra in the same class (i.e., an independent validation).

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Method
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\section{METHOD}
\label{sec: method}
 \subsection{Self Organizing Maps}
 \label{sec: som}

 The SOM is a clustering method which reduces the dimension of data to lower dimensions, usually 1 or 2D, while preserving topological features of the original data.
 Results of SOM contain nodes (neurons) that arranged in 1D or 2D arrays \citep{Kohonen98}.

 Each node may contain one or more samples from input data and distance between nodes represents similarity or dissimilarity of underlying samples. 
 In the way that similar data are closer together in the array and the further nodes go from each other, the more dissimilarity appears between their samples.
 A weight vector ``\boldit{W}" with same dimension of the input data associates with each node which will be changed during the process and has a key factor in a position of nodes in the map.

 \cite{Geach12} presented the application of the SOM and demonstrate the algorithm of the SOM in detail. In this section we are going to talk briefly about the algorithm of the SOM, how we create our maps and a test model which will help to interpret our results. 
 \subsubsection{Algorithm of SOM} 
 \label{sec: algorithm}
     Assuming we have a data set which contains vectors, \boldit{V} $\in \Re^n$, and we want to map them on S1 by S2 map. 
     We start by creating S1 $\times$ S2 empty neurons. 
     The initial arrangement of these neurons depends on a map's topology provided by users.
     In the case of the 1D maps, since each neuron has two immediate neighbours the topology of the map does not have any effect on the final result and we can choose any topology.
     However, in 2D maps, the shape of the neurons specifies the number of the immediate neighbours for each neuron and it is up to user that which shape is more suitable for their data.
     In this paper, we chose hexagonal topology which gives each neuron six neighbours, and causes more interactions between neurons.
     Then, a random weight vector, \boldit{W} $\in \Re^n$, will be assigned to each node.
     The process of creating SOM, happens over series of $N$ iterations. 
     During each iteration the weight vectors might change according to the Kohonen learning rule (equation~\ref{equ: weight adj}). 
      In each iteration SOM code:
     \begin{enumerate}
        \item Chooses a random vector from our data set ($V_i$).
        \item Calculates the Euclidean distance for each node, $j$, as  $D_j^2= \sum_{i=0}^{i=n} (V_i - W_i)^2$, and finds a neuron with minimum $D_j$, (``$D_{j_{min}}$"). This neuron is the winner node and is called the Best Matching Unit (BMU). 
        \item  Computes the radius of the neighbourhood of the BMU to find nodes within this radius. The weight vectors of these nodes will be affected in the next steps. This value is arbitrary and initially can be set to be as high as half of the SOM size and then it decades exponentially over each iteration:
        \begin{equation}
            r^t_{BMU} = r^0_{BMU}e^{(-t/\tau)}
        \end{equation}
        where $\tau$ is a decay constant and usually set to be the same as the number of iterations, $N$. $r^0_{BMU}$ and $r^t_{BMU}$ is the radius of the neighbourhood at 0th and $t$th iteration, respectively. 
        \item Changes the weight vectors of the BMU, and all the nodes within $r^t_{BMU}$ as:
        \begin{equation}
            \label{equ: weight adj}
            w(t+1)=w(t)+L(t) \times R(t) \times(v(t)-w(t))
        \end{equation}
        where $L(t) = L_0 e^{(-t/\tau)}$ is the learning factor which prevents divergence of the SOM and $R(t)=\exp(-\frac{D_j^2}{2r^t_{BMU}})$ is the influence rate. $R(t)$ determines how the weight of nodes in the neighbourhood of BMU will change.
        \item  Repeats these steps for $N$ times.
     \end{enumerate}

\subsection{Creating SOM}
\label{sec: create_som}
     In order to create SOM, we used {\tiny MATLAB} neural network 
     toolbox~\citep[NNT,][]{matlabtolbox}.
     %%Sr160611: Last time I got the following note from the MNRAS editor:
     %"Please note that computer software/programming lan- guages must be styled in SMALL CAPITAL LETTERS, according to journal style. Please check and correct this paper accordingly."
     SOM in {\tiny NNT} can be created by {\tiny NEWSOM} or {\tiny SELFORGMAP} library which both work in two phases. 
     Phase one is the ``ordering phase". 
     This phase starts with maximum neighbourhood distance, and initial high learning factor usually 0.9 which is provided by the user. 
     The ordering phase continues for the requested number of iterations.
     During the iterations, the learning factor reduces to tuning phase leaning factor and the neighbourhood distance reaches to tuning phase neighbourhood distance which both of the numbers set by the user.
     In the ordering phase, the changes in the learning factor and the neighbourhood distance adjust with the number of the iterations, in the way that at the last iteration, these two values reach to the initial values for the second phase.

     The second phase is the ``tuning phase".
     In this phase the neighbourhood distance is at its minimum, but learning factor decreases very slowly.
     This minimum neighbourhood distance and slowly decreasing the leaning factor helps to fine tune the topology results and causes the more stable SOM. 
     The number of iterations in this tuning phase most be much more than the number iterations in ordering phase, to allow the tuning to happen slowly. %cite kohenen book?!
     We chose the number of epochs the tuning phase to be 3 times more than the number of epochs in the ordering phase.

     To show our results, we used two of the {\tiny NNT}'s built-in plots.
     These plots, which are evolved version of old SOM plots, are designed to show distance between each clusters more clear.
     A Hits map, which shows the number of the times each neuron become the winner (hits), and a distance map, which shows the same neurons as the hits map and the distance between them.
     In the maps, the purple hexagonal shape areas represent the neurons.
     The distances in a distance map are shown by the grey cycle colours.
     The darker colour represents the larger distance between neurons; whilst the lighter colours means there is a small distance between neurons.
     In hit maps, neurons with zero hit left empty.

    Size of SOM maps is arbitrary and there is no rules or restrictions on choosing the size of the maps. 
    Although \cite{Vesanto05} suggested that the total number of  $5\sqrt{n}$ neurons provides the most sufficient size, users usually choose the size of grids based on their data set and their usages of the results.

\subsection{Mock sample}

         \begin{figure}
            \begin{subfigure}[b]{0.5\textwidth}
                \centering
                \includegraphics[width=\textwidth]{images0.01/sample/sample2_dist.png}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{0.5\textwidth}
                \includegraphics[width=\textwidth]{images0.01/sample/sample2_hits.png}
            \end{subfigure}
            \caption{SOM of the mock sample. In both plots, the axes show the position of the neurons. Hexagonal shapes represent the neurons. The upper plot is a distance map. The grey cycle colours show the differences between weight of each neuron with white is the minimum differences and black has the maximum one. The lower plot is a hit plot. It represents the number of samples in each neuron, while the empty with neurons means zero hit. In this sample the 27 galaxies clustered in 4 groups of 8, 7, 6 and 6 galaxies.}
            \label{fig: sample}
        \end{figure}

 We created  a mock sample of 27 galaxies to show how this method works, illustrate the results of SOMs, and how they can be analysed.
 The mock sample contains two information from each galaxy; The type of galaxies: type 1 or type 2, and either they are high or low star forming galaxies. 
 We generated a SOM with the size of $10 \times 10$ using the same initial values mentioned in the Section ~\ref{sec: create_som}.

 Fig. ~\ref{fig: sample} shows the SOM of the mock sample. 
 The upper plot is a distance map. 
 The axes show the position of the neurons in a $10 \times 10$ network.
 The lower plot in the Fig.~\ref{fig: sample} shows a hit map.
 On this map similar to the distance map the axes show the position of the neurons where the hexagonal shapes are the neurons.
 The purple neurons with a number on them shows the number of galaxies in them.
 Colour coverage of neurons is different and depends on the number of the hits on the sample.
 The neuron with the maximum number of hits is completely coloured while the empty neurons are white.

Using this method, as we predicted, we were able to divide the mock sample galaxies into 4 distinct groups: Type 1 with high SFR, type 1 with low SFR, type 2 with high SFR and type 2 with low SFR. 
The upper plot in the Fig.~\ref{fig: sample}, clearly shows this division.
In that plot, the upper part belongs to high star forming galaxies, while the lower part belongs to the low star forming galaxies.
The left part of the plot, is where type 1 galaxies belong to and the right side is for type 2 galaxies.
Grey to black colours shows the border between regions.
The lower plot in Fig.~\ref{fig: sample} shows only 4 neurons out of 100 were occupied. 
8 galaxies are type 1 galaxies with high SFR, 7 are type 2 galaxies with high SFR, 6 are type 1 galaxies with low SFR and the other 6 are type 2 galaxies with low SFR. 
Since the input data are so simple, each entry only could get either 0 or 1 as a type of galaxy and 0 or 0.5 as an indicator of  high or low SFR, only four of the neurons were filled.

We can conclude that although each galaxy had enough space to occupy any of the neurons in the network, because of the similarity of values they stayed only in four groups.
This network is considered as a train network, and if there are any data set with similar entries, we can use this network to cluster the new data set.

As we show in the following sections, in the real world with the real data, we never have two galaxies with exactly the same information, we have more than 2 dimensions, and more galaxy types. 
Therefore, if the network has high enough neurons, the input data would eventually separate from each other and cluster into smaller and smaller groups. 
However, if the information of the input data is very similar to each other, the number of neurons is going to be much higher than the number of input samples, to be able to separate the groups from each other. 
Therefore, it is up to the users that decide the similarity or dissimilarity between the input data based on number of neurons.

@@

    For each generated network, we compared the results with K96 categorization.
    We also used these networks to classify the T12 galaxies samples, and compared them with classifications from supervised networks in T12.
    \subsection{1D SOMs}
    \label{sec: 1D}
        \subsubsection{TRAINING THE NETWORKS}
        \label{sec: 1Dt}
            To start our clustering, we assumed that galaxies can be divided only in two general types; quiescent and starburst.
            In this case, we generated a network with only two neurons.
            We increased the size of the map gradually till we separated the input 12 sample galaxies into the 12 neurons.

            Figs.~\ref{fig: 1by2T} -~\ref{fig: 1by22T} show results of training networks.
            Similar to Fig.~\ref{fig: sample}, the upper part of the figures shows neurons and their relative distances, $D_j$, from each other.
            As it was mentioned in the Sec.~\ref{sec: method}, the colours between neurons represents the relative distance between the neurons and the darker is the colour the more relative distance is between them.
            The lower part shows how many galaxies from K96 template places in each neuron. 
            \begin{figure}
                \begin{subfigure}[b]{0.5\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{images0.01/1d/dist_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.5\textwidth}
                     \includegraphics[width=\textwidth]{images0.01/1d/hit_t_1_by_2.png}
                \end{subfigure}
                \caption{Results of training network in $1\times2$~grid. Similar to Fig.~\ref{fig: sample} the upper plot is a distance map and the lower plot is a hit map. In this network, 5 of galaxies from K96 model, categorized as quiescent galaxies and the rest are starbursts. Because of their high emission lines, Sc galaxies moved towards the starburst ones.}
                 \label{fig: 1by2T}
            \end{figure}

@@

            The fact that the galaxies had to have 22 different neurons till be divided into 12 groups, shows that the difference between SB1 and SB2, and B and E galaxies are very small.
            %These small differences might be very small.
            Therefore, they intend to stay in the same group till the network become big enough to make attention to the smallest particularity.
